---
title: "check_outputs_country"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
# library(tigris)
library(sf)
library(cowplot)
library(rebus)

## add paths
main_path <- '/Volumes/GoogleDrive/Shared Drives/emlab/projects/current-projects/climate-data-pipeline/'

save_path <- paste0(main_path, 'agg-outputs/diagnostic-figs/')

output_path <- paste0(main_path, 'agg-outputs/')

## files
weight_file  <- 'weights/eu_nuts2_era5_area_crop_weights.csv'

temp_file <- 'output/eu_area_era5_temp_average_1994_2010_polynomial_5_area_crop_weights.csv'

prcp_file <- 'output/eu_area_era5_prcp_sum_1994_2010_polynomial_3_area_crop_weights.csv'

polygon_file <- 'shapefiles/EU/NUTS2_RG_60M_2013_mortality.shp'


## Information for saving/filtering
## ----------------------------------------------

country_name <- "EU"
low_temp_month <- 'month_01'
high_temp_month <- 'month_08'
spring_months <- c('month_03', 'month_04', 'month_05')
fall_months <- c('month_10', 'month_11', 'month_12')
low_prcp_m <- 'month_12'
dry_months <- c('month_12', 'month_01', 'month_02')
high_prcp_m <- 'month_07'
wet_months <- c('month_06', 'month_07', 'month_08')
plot_yrs <- seq(1994, 2010, by = 5)




## Create folders as needed
  if(!dir.exists(paste0(save_path, country_name))){

    dir.create(paste0(save_path, country_name), recursive=T)

  }

if(!dir.exists(paste0(save_path, country_name, '/temp'))){

    dir.create(paste0(save_path, country_name, '/temp'), recursive=T)

}

if(!dir.exists(paste0(save_path, country_name, '/prcp'))){

    dir.create(paste0(save_path, country_name, '/prcp'), recursive=T)

  }


# Step 1 output - weights

# out_1 <- read_csv(file.path(here::here(), "data", "int", "weights", "ca_counties_era5_area_crop_weights.csv"))

out_1 <- read_csv(paste0(output_path, weight_file))


# Step 2 output
# out_2 <- read_csv(file.path(here::here(), "data", "output", "ca_counties_era5_temp_2005_2010_polynomial_3.csv"))

# Compare Step 2 output with crop weights 
# out_2_crop <- read_csv(file.path(here::here(), "data", "output", "ca_counties_era5_temp_2005_2010_polynomial_3_area_crop_weights.csv"))

out_2_crop_temp <- read_csv(paste0(output_path, temp_file))

# Compare Step 2 output with crop weights 

out_2_crop_prcp <- read_csv(paste0(output_path, prcp_file))



# # Untransformed monthly aggregated era5 data 
# era5_temp_daily <- read_csv(file.path(here::here(), "data", "output",
#                                    "ca_counties_test_temp_2010.csv"))
# 
# # Step 2 output precip 
# out_2_prcp <- read_csv(file.path(here::here(), "data", "output", "ca_counties_era5_prcp_sum_2005_2010_polynomial_3_area_crop_weights.csv"))
# 
# # Untransformed monthly era5 precip
# era5_prcp_daily <- read_csv(file.path(here::here(), "data", "output",
#                                    "ca_counties_test_prcp_2010.csv"))

# shapefile for checking maps
# us_counties <- tigris::counties()
# ca_counties <- us_counties %>% 
#   filter(STATEFP == '06')

country_shp <- read_sf(paste0(main_path, polygon_file)) 

## for plotting - change the column name as needed
country_areas <- country_shp %>% 
  mutate(poly_id = NUTS_ID) %>% 
  dplyr::select(poly_id)

if(country_name == "USA") {
  
  country_areas <- country_areas %>%
    mutate(poly_id = str_remove(poly_id, pattern = START %R% '0'))
  
}



```


# Check Step 1 Outputs

## Area Weights

Step 1 output is the overlap between polygons for Japan and the climate rasters. There should be `nrow(country_shp)` polygons, representing the `nrow(country_shp)` prefectures in the country. Area weights are the fraction of the polygon falling within each grid cell and are normalized by polygon so that all cells within a polygon sum to 1.  

```{r check-area-weights}
check_id_counts <- out_1 %>% 
  select(poly_id) %>% 
  distinct() 

nrow(country_shp) == nrow(check_id_counts)

# Map by polygon color 
map_polys <- ggplot() + 
  geom_raster(data = out_1, aes(x=x, y=y, fill=poly_id),
              alpha = 0.8) +
  # geom_sf(data = country_shp, color = 'black', fill = NA) +
  labs(x = '',
       y = '') + 
  theme_bw() + 
  theme(legend.position = '')

# Check x/y poly ids 
xy_ids <- out_1 %>% 
  group_by(x,y) %>% 
  summarize(n_poly_ids = n())

# Map area weights 
# Area weight = fraction of the cell covered by the polygon * area km 2
# Normalized by polygon so all polygons should sum to 1
map_area_weights <- 
  # country_shp %>%
  # st_shift_longitude() %>%
  ggplot() + 
  # geom_sf(fill = NA, lwd = 0) +
  geom_tile(data = out_1, aes(x=x, y=y, fill=w_area)) +
  labs(fill = 'Area Weight',
       x = '',
       y = '') + 
  scale_fill_viridis_c() + 
  # scale_x_continuous(limits = c(165, 184)) +
  theme_bw() 

# Check sum weights
sum_area_weights <- out_1 %>% 
  group_by(poly_id) %>% 
  mutate(sum_w_area = sum(w_area)) %>% 
  ungroup() # All = 1

unique(sum_area_weights$sum_w_area)




```

```{r message=F, warning=F, echo=F}
map_area_weights
```

## Crop Weights

Crop weights are the fraction of the crops within the polygon occurring within a given grid cell. These weights are also normalized by polygon so that all cells within a polygon sum to 1. 

```{r check-crop-weights}
# Map crop weights 
# crop weight = fraction of the cell covered by the polygon * crop weight
# Normalized by polygon so all polygons should sum to 1
map_crop_weights <- 
  # country_shp %>%
  # st_shift_longitude() %>%
  ggplot() + 
  geom_sf(fill = NA, lwd = 0) +
  geom_tile(data = out_1, aes(x=x, y=y, fill=weight)) +
  labs(fill = 'Area + Crop Weight',
       x = '',
       y = '') + 
  scale_fill_viridis_c() + 
  # scale_x_continuous(limits = c(165, 184)) +
  theme_bw() 

# Check sum weights
sum_crop_weights <- out_1 %>% 
  group_by(poly_id) %>% 
  mutate(sum_weight= sum(weight)) %>% 
  ungroup() # All = 1

unique(sum_crop_weights$sum_weight)

## save area and crop weights
weight_fig <- plot_grid(
  map_area_weights,
  map_crop_weights,
  align = "v",
  ncol = 1)

ggsave(weight_fig,
       filename = file.path(paste0(save_path, "/", country_name, '/weights_fig.pdf')),
       dpi = 300,
       device = 'pdf')


```

```{r message=F, warning=F, echo=F}
map_crop_weights
```

# Check Step 2 Output

## Temperature 

```{r}
# Check that each polygon has a value for all 12 months 
check_out2 <- out_2_crop_temp %>% 
  group_by(year, poly_id) %>% 
  summarize(n_months = n()) # all have 12

unique(check_out2$n_months)

# Check for any na values
check_nas <- out_2_crop_temp %>% 
  filter(is.na(order_1) | is.na(order_2) | is.na(order_3) | is.na(order_4) | is.na(order_5)) # None

nrow(check_nas)

# na_counties <- ca_counties %>% 
#   filter(GEOID %in% c('06015', '06023', '06025', '06071', '06073')) # San Diego, San Bernardino, Humboldt, Del Norte, Imperial 

# Convert the first order to avg temp by month so it's easier to understand the values 
# use value / days in month 

leap_years <- seq(1960, 2020, by = 4)


out_2_crop_temp2 <- out_2_crop_temp %>% 
  mutate(days_in_month = case_when(month %in% c('month_04', 'month_06', 'month_09', 'month_11') ~ 30,
                                   month == 'month_02' & year %in% leap_years ~ 29,
                                   month == 'month_02' & !year %in% leap_years ~ 28, 
                                   TRUE ~ 31)) %>% 
  mutate(order_1_avg = order_1 / days_in_month)

# # Check leap years
# feb <- out_2_crop_temp2 %>% 
#   filter(month == 'month_02')
# 
# leap_plot <- ggplot() + 
#   # geom_line(data = feb, aes(x = year, y = order_1, color = as.factor(poly_id), group = poly_id)) + 
#   geom_line(data = feb, aes(x = year, y = order_1_avg, color = as.factor(poly_id), group = poly_id)) + 
#   geom_vline(xintercept = 2016,
#              linetype = 'longdash') + 
#   scale_x_continuous(expand = c(0,0),
#                    breaks = seq(min(feb$year), max(feb$year), by=1)) + 
#   labs(x = 'Year',
#        y = 'Avg February Temperatures') + 
#   theme_bw() + 
#   theme(legend.position = '')

## Graphs 
# Polynomial values over time by polygon (one graph for each polynomial order)
poly1_plot <- ggplot() + 
  geom_line(data = out_2_crop_temp2, aes(x = month, y = order_1_avg, color = as.factor(poly_id), group = poly_id), alpha = 0.3) + 
  facet_wrap(~year) + 
  scale_x_discrete(expand = c(0,0),
                   breaks = c('month_01', 'month_03', 'month_05',
                              'month_07', 'month_09', 'month_11'),
                   labels = c('Jan', 'Mar', 'May', 'Jul',
                              'Sept', 'Nov')) +
  scale_y_continuous(expand = c(0,0),
                     limits = c(-40,40),
                     breaks = seq(-10,40,10)) +
  labs(x = 'Month',
       y = 'Averarge Monthly Temp by Polygon\n(First Order / Days in Month)') + 
  theme_bw() + 
  theme(legend.position = '')

ggsave(poly1_plot, filename = paste0(save_path, country_name, '/temp/poly1_temp.pdf'))

poly2_plot <- ggplot() + 
  geom_line(data = out_2_crop_temp2, aes(x = month, y = order_2, color = as.factor(poly_id), group = poly_id), alpha = 0.3) + 
  facet_wrap(~year) + 
  scale_x_discrete(expand = c(0,0),
                   breaks = c('month_01', 'month_03', 'month_05',
                              'month_07', 'month_09', 'month_11'),
                   labels = c('Jan', 'Mar', 'May', 'Jul',
                              'Sept', 'Nov')) + 
  scale_y_continuous(expand = c(0,0)) +
  labs(x = 'Month',
       y = 'Summed Monthly Temperature (Second Order)\nby Polygon') + 
  theme_bw() + 
  theme(legend.position = '')

ggsave(poly2_plot, filename = paste0(save_path, country_name, '/temp/poly2_temp.pdf'))


poly3_plot <- ggplot() + 
  geom_line(data = out_2_crop_temp2, aes(x = month, y = order_3, color = as.factor(poly_id), group = poly_id), alpha = 0.3) + 
  facet_wrap(~year) + 
  scale_x_discrete(expand = c(0,0),
                   breaks = c('month_01', 'month_03', 'month_05',
                              'month_07', 'month_09', 'month_11'),
                   labels = c('Jan', 'Mar', 'May', 'Jul',
                              'Sept', 'Nov')) + 
  scale_y_continuous(expand = c(0,0),
                     labels = scales::comma) +
  labs(x = 'Month',
       y = 'Summed Monthly Temperature (Third Order)\n by Polygon') + 
  theme_bw() + 
  theme(legend.position = '')

ggsave(poly3_plot, filename = paste0(save_path, country_name, '/temp/poly3_temp.pdf'))


## maps
## --------------------------------------------------

# country_check <- country_areas %>%
#   st_shift_longitude() %>%
#   ggplot() +
#   geom_sf(aes(), fill = "navy", lwd = 0) +
#   labs(x = '',
#        y = '') +
#   theme_bw()
# 
# ggsave(country_check, filename = paste0(save_path, country_name, '/country.pdf'))


out_3 <- out_2_crop_temp2 %>% 
  filter(!is.na(month))

out3_sf <- merge(country_areas, out_3, all = TRUE)

# Map first order values by region for low temp month
low_temp_df <- out3_sf %>%
  filter(month == low_temp_month)

map_order1_lowt <- low_temp_df %>% 
  st_shift_longitude() %>%
  filter(year %in% plot_yrs) %>%
  ggplot() + 
  geom_sf(aes(fill = order_1_avg), lwd = 0) + 
  labs(fill = paste0('Avg ', low_temp_month, '\nTemperature\n(1st Order / Days in Month)'),
       title = 'area + crop weights',
       x = '',
       y = '') + 
  scale_fill_viridis_c(limits = c(round(min(low_temp_df$order_1_avg)) - 1, round(max(low_temp_df$order_1_avg)) + 1)) + 
  facet_wrap(~year) + 
  theme_bw() 

ggsave(map_order1_lowt, filename = paste0(save_path, country_name, '/temp/low_temp_out.pdf'),
       width = 11,
       height = 10,
       units = "in")

# Map first order values for high temp month (July)
high_temp_df <-  out3_sf %>%
  filter(month == high_temp_month)

map_order1_hight <- high_temp_df %>%
  st_shift_longitude() %>%
  filter(year %in% plot_yrs) %>%
  ggplot() + 
  geom_sf(aes(fill = order_1_avg), lwd = 0) + 
  labs(fill = paste0('Avg ', high_temp_month, '\nTemperature\n(1st Order / Days in Month)'),
       title = 'area + crop weights',
       x = '',
       y = '') + 
  scale_fill_viridis_c(limits = c(round(min(high_temp_df$order_1_avg)) - 1, round(max(high_temp_df$order_1_avg)) + 1)) + 
  facet_wrap(~year) + 
  theme_bw() 

ggsave(map_order1_hight, filename = paste0(save_path, country_name, '/temp/high_temp_out.pdf'),
       width = 11,
       height = 10,
       units = "in")

## map first order for three consecutive months 
spring_df <-  out3_sf %>%
  filter(month %in% spring_months) 

map_order1_spring <- spring_df %>%
  st_shift_longitude() %>%
  filter(year %in% plot_yrs) %>%
  ggplot() + 
  geom_sf(aes(fill = order_1_avg), lwd = 0) + 
  labs(fill = 'Avg spring\nTemperature\n(1st Order / Days in Month)',
       title = 'area + crop weights',
       x = '',
       y = '') + 
  scale_fill_viridis_c(limits = c(round(min(spring_df$order_1_avg)) - 1, round(max(spring_df$order_1_avg)) + 1)) + 
  facet_grid(year~month) + 
  theme_bw() 

ggsave(map_order1_spring, filename = paste0(save_path, country_name, '/temp/spring_temp_out.pdf'),
       width = 11,
       height = 10,
       units = "in")

## map first order for three consecutive months (sept, oct, nov)
fall_temp_df <- out3_sf %>% 
  filter(month %in% fall_months)

map_order1_fall <- fall_temp_df %>%
  st_shift_longitude() %>%
  filter(year %in% plot_yrs) %>%
  ggplot() + 
  geom_sf(aes(fill = order_1_avg), lwd = 0) + 
  labs(fill = 'Avg fall\nTemperature\n(1st Order / Days in Month)',
       title = 'area + crop weights',
       x = '',
       y = '') + 
  scale_fill_viridis_c(limits = c(round(min(fall_temp_df$order_1_avg)) - 1, round(max(fall_temp_df$order_1_avg)) + 1)) +  
  facet_grid(year~month) + 
  theme_bw() 

ggsave(map_order1_fall, filename = paste0(save_path, country_name, '/temp/fall_temp_out.pdf'),
       width = 11,
       height = 10,
       units = "in")

# ## Maps for output with crop weights 
# # Convert the first order to avg temp by month so it's easier to understand the values 
# out_2_crop <- out_2_crop %>% 
#   mutate(days_in_month = case_when(month %in% c('month_04', 'month_06', 'month_09', 'month_11') ~ 30,
#                                    month == 'month_02' & year == 2008 ~ 29,
#                                    month == 'month_02' & year != 2008 ~ 28, 
#                                    TRUE ~ 31)) %>% 
#   mutate(order_1_avg = order_1 / days_in_month)
# 
# out2_crop_sf <- merge(ca_counties, out_2_crop, all = TRUE)
# 
# # Map first order values by polygon for lowest temp month (Feb) 
# map_order1_feb_crop <- out2_crop_sf %>% 
#   filter(month == 'month_02') %>% 
#   ggplot() + 
#   geom_sf(aes(fill = order_1_avg)) + 
#   labs(fill = 'Avg Feb\nTemperature\n(1st Order)',
#        title = 'area + crop weights',
#        x = '',
#        y = '') + 
#   scale_fill_viridis_c(limits = c(-7, 37)) + 
#   facet_wrap(~year) + 
#   theme_bw() 
# 
# # Map first order values for highest temp month (July)
# map_order1_jul_crop <- out2_crop_sf %>% 
#   filter(month == 'month_07') %>% 
#   ggplot() + 
#   geom_sf(aes(fill = order_1_avg)) + 
#   labs(fill = 'Avg July\nTemperature\n(1st Order)',
#        title = 'area + crop weights',
#        x = '',
#        y = '') + 
#   scale_fill_viridis_c(limits = c(-7, 37)) + 
#   facet_wrap(~year) + 
#   theme_bw() 
```

```{r message=F, warning=F, echo=F}
map_order1_feb
map_order1_feb_crop
```


```{r message=F, warning=F, echo=F}
map_order1_jul
map_order1_jul_crop
```

## Precipitation

```{r}
# Check that each polygon has a value for all 12 months 
check_out2 <- out_2_crop_prcp %>% 
  group_by(year, poly_id) %>% 
  summarize(n_months = n()) # Looks good

unique(check_out2$n_months)

# ## convert from summation
# out_2_crop_prcp2 <- out_2_crop_prcp %>% 
#   mutate(days_in_month = case_when(month %in% c('month_04', 'month_06', 'month_09', 'month_11') ~ 30,
#                                    month == 'month_02' & year %in% leap_years ~ 29,
#                                    month == 'month_02' & !year %in% leap_years ~ 28, 
#                                    TRUE ~ 31)) %>% 
#   mutate(order_1_avg = order_1 / days_in_month)

## Maps 
# Join first order outputs to the shp 

out2_prcp_sf <- merge(country_areas, out_2_crop_prcp, all = TRUE)

# Map first order values by area for lowest precip month (december) 
low_prcp <- out2_prcp_sf %>% 
  filter(month == low_prcp_m)

map_prcp_low <- low_prcp %>% 
  st_shift_longitude() %>%
  filter(year %in% plot_yrs) %>%
  ggplot() + 
  geom_sf(aes(fill = order_1), lwd = 0) + 
  labs(fill = paste0('Total ', low_prcp_m, '\nPrecipitation (mm)\n(1st Order)'),
       title = 'area + crop weights',
       x = '',
       y = '') + 
  scale_fill_viridis_c(option = 'mako', direction = -1,
                       limits = c(round(min(low_prcp$order_1)) - 1, round(max(low_prcp$order_1)) + 1)) + 
  facet_wrap(~year) + 
  theme_bw() 

ggsave(map_prcp_low, filename = paste0(save_path, country_name, '/prcp/low_prcp_out.pdf'),
       width = 11,
       height = 10,
       units = "in")


# Map first order values for highest precip month (June)
high_prcp <- out2_prcp_sf %>% 
  filter(month == high_prcp_m)

map_prcp_high <- high_prcp %>% 
  st_shift_longitude() %>%
  filter(year %in% plot_yrs) %>%
  ggplot() + 
  geom_sf(aes(fill = order_1), lwd = 0) + 
  labs(fill = 'Total ', high_prcp_m, '\nPrecipitation (mm)\n(1st Order)',
       title = 'area + crop weights',
       x = '',
       y = '') + 
  scale_fill_viridis_c(option = 'mako', direction = -1,
                       limits = c(round(min(high_prcp$order_1)) - 1, round(max(high_prcp$order_1)) + 1)) + 
  facet_wrap(~year) + 
  theme_bw() 

ggsave(map_prcp_high, filename = paste0(save_path, country_name, '/prcp/high_prcp_out.pdf'),
       width = 11,
       height = 10,
       units = "in")

## map first order for three consecutive wet months 
wet_months <- out2_prcp_sf %>% 
  filter(month %in% wet_months)

map_order1_wet <- wet_months %>% 
  st_shift_longitude() %>%
  filter(year %in% plot_yrs) %>%
  ggplot() + 
  geom_sf(aes(fill = order_1), lwd = 0) + 
  labs(fill = 'Monthly Precipitation (mm)\nWet Season\n(1st Order)',
       title = 'area + crop weights',
       x = '',
       y = '') + 
  scale_fill_viridis_c(option = 'mako', direction = -1,
                       limits = c(round(min(wet_months$order_1)) - 1, round(max(wet_months$order_1)) + 1)) +  
  facet_grid(year~month) + 
  theme_bw() 

ggsave(map_order1_wet, filename = paste0(save_path, country_name, '/prcp/wet_season_out.pdf'),
       width = 11,
       height = 10,
       units = "in")

## map first order for three consecutive months 
dry_months <- out2_prcp_sf %>% 
  filter(month %in% dry_months)

map_order1_dry <- dry_months %>% 
  st_shift_longitude() %>%
  filter(year %in% plot_yrs) %>%
  ggplot() + 
  geom_sf(aes(fill = order_1), lwd = 0) + 
  labs(fill = 'Total Precipitation (mm)\nDry Season\n(1st Order)',
       title = 'area + crop weights',
       x = '',
       y = '') + 
  scale_fill_viridis_c(option = 'mako', direction = -1,
                       limits = c(round(min(dry_months$order_1)) - 1, round(max(dry_months$order_1)) + 1)) +  
  facet_grid(year~month) + 
  theme_bw() 

ggsave(map_order1_dry, filename = paste0(save_path, country_name, '/prcp/dry_season_out.pdf'),
       width = 11,
       height = 10,
       units = "in")

## plot first, second, third order

# Polynomial values over time by polygon (one graph for each polynomial order)
poly1_plot_p <- ggplot() + 
  geom_line(data = out_2_crop_prcp, aes(x = month, y = order_1, color = as.factor(poly_id), group = poly_id), alpha = 0.3) + 
  facet_wrap(~year) + 
  scale_x_discrete(expand = c(0,0),
                   breaks = c('month_01', 'month_03', 'month_05',
                              'month_07', 'month_09', 'month_11'),
                   labels = c('Jan', 'Mar', 'May', 'Jul',
                              'Sept', 'Nov')) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = 'Month',
       y = 'Total Monthly Prcp by Polygon (mm)\n(First Order)') + 
  theme_bw() + 
  theme(legend.position = '')

ggsave(poly1_plot_p, filename = paste0(save_path, country_name, '/prcp/poly1_prcp.pdf'))

# Polynomial values over time by polygon (one graph for each polynomial order)
poly2_plot_p <- ggplot() + 
  geom_line(data = out_2_crop_prcp, aes(x = month, y = order_2, color = as.factor(poly_id), group = poly_id), alpha = 0.3) + 
  facet_wrap(~year) + 
  scale_x_discrete(expand = c(0,0),
                   breaks = c('month_01', 'month_03', 'month_05',
                              'month_07', 'month_09', 'month_11'),
                   labels = c('Jan', 'Mar', 'May', 'Jul',
                              'Sept', 'Nov')) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = 'Month',
       y = 'Total Monthly Prcp by Polygon\n(Second Order)') + 
  theme_bw() + 
  theme(legend.position = '')

ggsave(poly2_plot_p, filename = paste0(save_path, country_name, '/prcp/poly2_prcp.pdf'))

poly3_plot_p <- ggplot() + 
  geom_line(data = out_2_crop_prcp, aes(x = month, y = order_3, color = as.factor(poly_id), group = poly_id), alpha = 0.3) + 
  facet_wrap(~year) + 
  scale_x_discrete(expand = c(0,0),
                   breaks = c('month_01', 'month_03', 'month_05',
                              'month_07', 'month_09', 'month_11'),
                   labels = c('Jan', 'Mar', 'May', 'Jul',
                              'Sept', 'Nov')) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = 'Month',
       y = 'Total Monthly Prcp by Polygon\n(Third Order)') + 
  theme_bw() + 
  theme(legend.position = '')

ggsave(poly3_plot_p, filename = paste0(save_path, country_name, '/prcp/poly3_prcp.pdf'))




```


# Compare to Raw Data

## Temperature 

Compare the above Feb & July maps to the raw era5 data aggregated to the monthly scale for 2010. These data haven't gone through any of the non-linear transformations. 

```{r}
# Back to -180 to 180
# Add days in the month 
era5_temp_daily <- era5_temp_daily %>% 
  mutate(x = ifelse(x>180, x-360, x),
         days_in_month = case_when(month %in% c('month_04', 'month_06', 'month_09', 'month_11') ~ 30,
                                   month == 'month_02' ~ 28, 
                                   TRUE ~ 31))

era5_t_monthly_by_grid <- era5_temp_daily %>% 
  group_by(x, y, poly_id, month, days_in_month) %>% 
  summarize(sum_temp = sum(value)) %>% 
  mutate(avg_temp = sum_temp / days_in_month)

## Maps for untransformed era5 data 
# February 
era5_temp_feb <- era5_t_monthly_by_grid %>% 
  filter(month == 'month_02')

map_era5_t_feb <- ggplot() + 
  geom_raster(data = era5_temp_feb,
              aes(x=x, y=y, fill = avg_temp)) + 
  geom_sf(data = ca_counties, fill=NA, color='black') + 
  labs(fill = 'Avg Feb\nTemperature\n(Raw ERA5)',
       title = 'Montly ERA5 (2010)',
       x = '',
       y = '') + 
  scale_fill_viridis_c(limits = c(-7, 37)) + 
  theme_bw() 

# July 
era5_temp_jul <- era5_monthly_by_grid %>% 
  filter(month == 'month_07')

map_era5_t_jul <- ggplot() + 
  geom_raster(data = era5_temp_jul,
              aes(x=x, y=y, fill = avg_temp)) + 
  geom_sf(data = ca_counties, fill=NA, color='black') + 
  labs(fill = 'Avg July\nTemperature\n(Raw ERA5)',
       title = 'Montly ERA5 (2010)',
       x = '',
       y = '') + 
  scale_fill_viridis_c(limits = c(-7, 37)) + 
  theme_bw() 
```

## Precipitation 

Compare the above July & December maps to the raw era5 data aggregated to the monthly scale for 2010. These data haven't gone through any of the non-linear transformations. 

```{r}
# Back to -180 to 180
# Add days in the month 
era5_prcp_daily <- era5_prcp_daily %>% 
  mutate(x = ifelse(x>180, x-360, x))

era5_p_monthly_by_grid <- era5_prcp_daily %>% 
  group_by(x, y, poly_id, month) %>% 
  summarize(sum_prcp = sum(value)) 

## Maps for untransformed era5 data 
# July
era5_prcp_jul <- era5_p_monthly_by_grid %>% 
  filter(month == 'month_07')

map_era5_p_jul <- ggplot() + 
  geom_raster(data = era5_prcp_jul,
              aes(x=x, y=y, fill = sum_prcp)) + 
  geom_sf(data = ca_counties, fill=NA, color='black') + 
  labs(fill = 'Total July\nPrecipitation (meters)',
       title = 'Montly ERA5 (2010)',
       x = '',
       y = '') + 
  scale_fill_viridis_c(option = 'mako', direction = -1,
                       limits = c(0, 0.07)) + 
  theme_bw() 

# December 
era5_prcp_dec <- era5_p_monthly_by_grid %>% 
  filter(month == 'month_12')

map_era5_p_dec <- ggplot() + 
  geom_raster(data = era5_prcp_dec,
              aes(x=x, y=y, fill = sum_prcp)) + 
  geom_sf(data = ca_counties, fill=NA, color='black') + 
  labs(fill = 'Total Dec\nPrecipitation (meters)',
       title = 'Montly ERA5 (2010)',
       x = '',
       y = '') + 
  scale_fill_viridis_c(option = 'mako', direction = -1,
                       limits = c(0, 0.61)) + 
  theme_bw() 
```